Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-11-17 special-issue-ci/Response to reviewers.tex

We have added a sentence clarifying that our intention was simply to
point out that the pattern is similar. We are obviously in a wholly
different league.

\section{Response to Reviewer \#4}

\begin{quote}
This paper presents a very challenging proposal. Namely, authors introduce a novel client-server
architecture for low cost and "potentially" massive distributed evolutionary algorithms. The main novelty
turns up from the fact that the architecture is developed in JavaScript.
However, JavaScript is well-known because of its lack of capability to deal with high performance
applications like the ones that are addressed in this paper. Thus, the use of JavaScript makes hard to
believe in the possibility of "potentially" massive computation as
sketched by the authors.
\end{quote}

JavaScript is not, in fact, so slow and to prove its speed we have
published a technical report comparing it with other compiled
languages (such as Scala) and scripted languages
\cite{2015arXiv151101088M}. In fact, node.js (used for the benchmarks)
and the JavaScript virtual machine in Chrome (used in Android phones
and Chrome and Chromium browsers) share the same JIT compiler, so
their speeds are virtually the same. In general, as a conclusion of
the report we see that the real difference between scripted and
compiled languages will vary with the language, but also with the
particular implementation of the algorithm and the size of the problem
or chromosome you are using. Even if the speed for a particular
problem is two or five times that achieved for JavaScript, this
disadvantage is more than offset by the availability of that free
resource in abundance; that is, you will only need that many
volunteers cooperating to get better performance than what you would
get with C or Java. Even so, this NodIO architecture, as highlighted
in the new version of the paper, is pool-based so local sources using
high-speed devices such as GPUs or even local powerful CPUs can be
added as non-volunteer users, since the heterogeinity and simple API
of the system allows that by design.

@@

\begin{quote}
The paper is well written and organized. State of the art is quite complete. Nevertheless, the experimental
section must be carefully checked and extended in order to fully validate the goodness of the proposed
architecture. At least, authors should show the behavior of the architecture in a real-world problem
demanding high computational performance.
Finally, English must be double-checked in order to correct minor typos such as "in the sense than[that]"
(in page 2).
\end{quote}

We have extended the experiments section with a new one, based on the
hard Rastrigint problem.

### 2015-11-17 special-issue-ci/Response to reviewers.tex

We have also justified our {\em pragmatic} approach to security
increasing trust by the user, but also monitoring their behavior to
detect it and eventually ban users. So far, this has not happened.

\begin{quote{
2) Regarding the experiments I am very skeptical about their real value. You chose the "Trap"
funcion. in which setting all bits to zero gives local maximum and setting all bits to one gives global
maximum. In the experiments you counted how many times evolutionary algorithm was able to find
global maximum for the problem of size n=40. Initially you have found that if the population in an
evolutionary algorithm is bit enough (1024 individuals) the algorithm returned global maximum
within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run
(page 4, if I understood it correctly).
\end{quote}
I fact 3.46 is the average for all 50 runs. We have enphasized the
word average to make that clearer.

\begin{quote}
Thus, what was the point of using distributed computing to solve such simple problem? Why the
authors did not perform their experiments for larger values of n, what could justify distributed problem
solution. You admit that the NodIO system rarely beats a high-performance desktop computer (page
6). Situation is better for the NodIO-W2 using JS Web Workers and automatic restarts, nevertheless,
it is still behind the desktop computer in more than 60% cases (when compared to the second
baseline). I suggest to perform additional computing experiments for some higher value of n in the
"Trap" function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
to focus the paper on the behavior analysis of volunteers participating in a distributed computing
system, what may be also interesting an valuable for the potential
readers.
\end{quote}

In fact, that was never the point of the paper. We have chosen Trap as
a relatively difficult problem that allows us mainly to test scaling
and measure behavior. In a production environment, a volunteer-only
environment would only be used if there is absolutely no other
resource available; in general, it would be combined with the rest of
the resources available to the experimenter: cloud or local; it would
make no sense to not use them if they are available and in fact, NodIO
allows to combine them seamlessly in a single system. That has been
explained a bit better in the introduction and conclusions
section. Besides, we have added a more difficult problem, the shifted
Rastrigin function, to make measures in a different scenario (one in
which the problem is not solved in a few seconds).

### 2015-11-17 special-issue-ci/Response to reviewers.tex

We have also justified our {\em pragmatic} approach to security
increasing trust by the user, but also monitoring their behavior to
detect it and eventually ban users. So far, this has not happened.

\begin{quote{
2) Regarding the experiments I am very skeptical about their real value. You chose the "Trap"
funcion. in which setting all bits to zero gives local maximum and setting all bits to one gives global
maximum. In the experiments you counted how many times evolutionary algorithm was able to find
global maximum for the problem of size n=40. Initially you have found that if the population in an
evolutionary algorithm is bit enough (1024 individuals) the algorithm returned global maximum
within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run
(page 4, if I understood it correctly).
\end{quote}
I fact 3.46 is the average for all 50 runs. We have enphasized the
word average to make that clearer.

\begin{quote}
Thus, what was the point of using distributed computing to solve such simple problem? Why the
authors did not perform their experiments for larger values of n, what could justify distributed problem
solution. You admit that the NodIO system rarely beats a high-performance desktop computer (page
6). Situation is better for the NodIO-W2 using JS Web Workers and automatic restarts, nevertheless,
it is still behind the desktop computer in more than 60% cases (when compared to the second
baseline). I suggest to perform additional computing experiments for some higher value of n in the
"Trap" function or to choose another type of problem like some NP-hard problem. Otherwise I suggest
to focus the paper on the behavior analysis of volunteers participating in a distributed computing
system, what may be also interesting an valuable for the potential
readers.
\end{quote}

In fact, that was never the point of the paper. We have chosen Trap as
a relatively difficult problem that allows us mainly to test scaling
and measure behavior. In a production environment, a volunteer-only
environment would only be used if there is absolutely no other
resource available; in general, it would be combined with the rest of
the resources available to the experimenter: cloud or local; it would
make no sense to not use them if they are available and in fact, NodIO
allows to combine them seamlessly in a single system. That has been
explained a bit better in the introduction and conclusions
section. Besides, we have added a more difficult problem, the shifted
Rastrigin function (a well known function optimization problem), to make 
measures in a different scenario (one in which the problem is not solved 
in a few seconds).

### 2015-11-17 special-issue-ci/Response to reviewers.tex

 - Safety - in my opinion is the biggest concern in such kind of distributed systems. First of all the
 code of evolutionary algorithm is sent to the client as a source code in Java Script, so some
 experienced user may quite simple decode what has to be computed. It can may come into
 possession of sensitive data or/and alter this code in order to disrupt the computations or make
 them worhless. The results send to the server may also be captured and/or altered without any
 problem since the transmission is not secure.

 In my opinion you should at least highlight above issues.
\end{quote}

We have also justified our {\em pragmatic} approach to security
increasing trust by the user, but also monitoring their behavior to
detect it and eventually ban users. So far, this has not happened.

\begin{quote{
2) Regarding the experiments I am very skeptical about their real value. You chose the "Trap"
funcion. in which setting all bits to zero gives local maximum and setting all bits to one gives global
maximum. In the experiments you counted how many times evolutionary algorithm was able to find
global maximum for the problem of size n=40. Initially you have found that if the population in an
evolutionary algorithm is bit enough (1024 individuals) the algorithm returned global maximum
within 3.46 sec. divided by 50 runs - i.e. 0.692 sec. for a single run
(page 4, if I understood it correctly).
\end{quote}

I fact 3.46 is the average for all 50 runs. 
We have emphasized the word average to make that clearer.

### 2015-11-17 special-issue-ci/Response to reviewers.tex

\begin{quote}

5. Length: As I mentioned above the authors should think of removing some of the unnecessary graphs
in order to improve readability. They admit themselves that it is hard to conclude from some of the
graphs (Fig. 5 and Fig. 8). Nevertheless, overall length of the paper is fine.

\end{quote}

Finally, since the reviewer thinks that length is fine and for the
reasons outlined above, we have decided to keep the graphs and explain
them a bit better.

@@

\begin{quote}
6. Abstract: Abstract is adequate, moreover it puts an accent on the analysis of volunteers' behavior in a
distributed computing system, contrary to the description in the main text of the paper. However, it
lacks the information about the second version of the system called NodIO-W2 based on the Web
Worker architecture.
\end{quote}

We have extended the abstract stressing our main points, that were
first to test the validity and measure the performance of our
framework NodIO and then to use the measures as first steps to a model
of volunteer uses.

@@

1) I like the state of the art which includes comprehensive overview of the most important papers related
to the subject presented in the paper. I also like the description of the system, but I cannot fully agree
with its characteristic provided on page 4. Namely:

 - Heterogeneity - it is true that almost all devices can participate in the web browser based
 distributed computing system, providing that a user did not disable Java Script in its browser.
 However if the authors plan to use HTML5 WebWorkers only modern web browsers can
 participate in such computing system, and some devices (e.g. based on Android 2.2-4.3 or
 Windows Phone) are also excluded.

\end{quote}

That is true, but it makes it anyway the most inclusive system so
far, as we have tested it with smartphones, tablets and computers. In
general, nowadays with rich internet applications disabling JavaScript
is not common, but anyway we do not aim for {\em all} clients, we have
enough with {\em most}. Windows Phone, for instance and maybe
fortunately, is not too common. % ¿no existe Chrome para Windows Phone? De todas formas yo quitaría lo de "and maybe fortunately"...

### 2015-11-17 special-issue-ci/volunteer-clouds-ieee-ci-si.tex
fixing a typo in the paper

In the second implementation, {\sf NodIO$^2$}, Web Workers
were used so that several clients per browser could be executed and
they could run in the background, so that the the number of
generations per user could be improved. The number of generations per
user increased to the point that baseline performance was improved in
most cases. The architectural changes led to a high correlation
between successive experiments, so that, even as the median number of
volunteers per experiment decreased, they were probably contributing
to the experiment at the same time, so that our objective was
achieved. In general, one of the first conclusion for this paper is
that the browser technologies should be used to its full extent so
that the user time is leveraged to its full potential, and this was
achieved with {\sf NodIO$^2$}.

>>>>

In the second implementation, {\sf NodIO$^2$}, Web Workers
were used so that several clients per browser could be executed and
they could run in the background, so that the the number of
generations per user could be improved. The number of generations per
user increased to the point that baseline performance was improved in
most cases. The architectural changes led to a high correlation
between successive experiments, so that, even as the median number of
volunteers per experiment decreased, they were probably contributing
to the experiment at the same time, so that our objective was
achieved. In general, one of the first conclusions for this paper is
that the browser technologies should be used to its full extent so
that the user time is leveraged to its full potential, and this was
achieved with {\sf NodIO$^2$}.
